The Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams

### 2016-09-20 apjRevised/response.txt

REVIEWER:
The authors investigate the linear kinetic instability of ultrarelativistic pairs against the excitation of oblique electrostatic oscillations in the unmagnetized
intergalactic medium, where the ultrarelativistic pairs result from the annihilation of TeV photons from extragalactic blazars with the extragactic backgound light (EBL). They demonstrate that the linear growth times of the instability are substantially shorter than the inverse Compton (with the cosmic microwave background radiation (CMB)) energy loss time of the pairs, and conclude from that that most of the kinetic energy of the pairs is dissipated into the electrostatic turbulence.

In the literature only the linear kinetic instability for electrostatic oscillations with wave vectors parallel to the beam direction has been calculated (Schlickeiser, Krakau and Supsar 2013, ApJ 777, 49), whereas oblique electrostatic oscillations were only treated in the reactive limit (Schlickeiser, Ibscher and Supsar 2012, ApJ
758, 102). Therefore the analysis of the authors in principle could be a valuable addition to the literature. However, for two reasons, detailed below, the manuscript in its present form is not acceptable for publication.

RESPONSE:

We thank the referee for a thoughtful report.

REVIEWER:

1) The authors calculated the linear growth rate by adopting the inappropriate relativistically drifting Maxwellian distribition function (Eqs. (1) and (3)) for the pair
production spectrum from the double photon annihilation process. But this Maxwellian distribution function does not result from the double photon annihilation process. The correct pair production spectrum has been calculated before in Eq. (154) of Schlickeiser et al. (2012, ApJ 758, 101) -- see also the approximations (3) and (5) in Schlickeiser, Krakau and Supsar (2013, ApJ 777, 49) -- and by Aharonian, Atoyan and Nagapetian (1983, Astrofizika (Astrophysics), 19, 1), depending on the incoming TeV gamma-ray enery spectrum and the EBL target photon spectrum. This results in a power law energy spectrum with a strong exponential cut-off at small electron energies, which is substantially different from a drifting Maxwellian. It is necessary for the authors to redo their linear instability analysis with the correct pair production energy spectrum.

RESPONSE:

We agree that a drifting Maxwellian is not the correct distribution function to model a pair beam. But we adopt it because:
1. It is simpler and allows us to calculate the kinetic instability exactly in the electrostatic approximation.
2. It has all the features that allows us the elucidate the physics (continuous spread of momenta in the parallel and perpendicular directions)
3. It is relatively not ad-hoc — by this we mean that it is a distribution that can be realized in physical situation.

We have added an additional paragraph to the end of section 2.1 elucidating this choice and saying that a correct calculation involves using a distribution function calculated from the \gamma\gamma process such as the ones mentioned by the referee.  We also note that the analytic techniques that we used for calculating the drifting Maxwellian may be useful in the full calculation.

REVIEWER:

2) Claiming that most of the kinetic energy of the pairs is dissipated into electrostatic turbulence on the ground, that the minimum linear instability groth time is shorter
than the inverse Compton energy loss time, is insufficient and not correct. The backreaction of the electrostatic turbulence on the unstable pair distribution function will ultimately lead to a plateaued distribution function in the original beam direction (see Menzler and Schlickeiser 2015, MNRAS 448, 3405) on a time scale much longer than the minimum linear growth time. The pairs are so redistributed in energy but still able to inverse Compton scatter the CMB photons. The authors should address this redistribution time scale and compare it with the inverse Compton energy loss time scale.

RESPONSE:

We completely agree with the referee that this statement is not correct and while we do not make this claim in this paper, we can understand how the discussion in the introduction can give this impression.  We were aware of these subtleties and focus just on the linear physics in this paper avoid any quantitative comparison with the ICC timescales.  We merely confirmed the linear results uses in earlier papers, clarified the physics, especially in regard to the broad support of modes that are unstable, and address the transition between the kinetic and reactive mode for the oblique instability.

To help clarify this discussion, we have included in the 4th paragraph of the introduction:

We note that this effectiveness of this plasma instabilities is complicated by nonlinear effects, which we briefly discuss below.

We have also added a paragraph below the sixth paragraph

### 2016-09-20 apjRevised/ms.tex
revision response

%\documentclass[usenatbib,iop,apj]{emulateapj}
\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2015}

@@

\usepackage[normalem]{ulem}
\usepackage[usenames]{color}

\definecolor{DarkGreen}{rgb}{0.64,0.80,0.35}
\newcommand\ac[1]{{\color{green} \bf #1}} %% AEB color
\newcommand\ab[1]{{\color{green} \bf #1}} %% AEB color
\newcommand\cp[1]{{\color{blue} #1}} %% CP color
\newcommand\ep[1]{{\color{magenta} \bf #1}} %% EP color
\newcommand\asout[1]{\color{green}\sout{{\rm  #1}}} %% AEB strikeout of text
\newcommand\qc[1]{{\color{red} \bf #1}}
%\newcommand\qc[1]{{\color{black} #1}}
\newcommand\cc[1]{{\color{blue} \bf #1}}
\newcommand\phil[1]{{\color{cyan} #1}}
\newcommand{\mathsc}[1]{\text{\textsc{#1}}}

@@

%\newcommand{\vel}{\ensuremath{\boldsymbol v}}
%\newcommand{\pmom}{\ensuremath{\boldsymbol p}}
%\newcommand{\Efield}{\ensuremath{\boldsymbol E}}
%\newcommand{\Bfield}{\ensuremath{\boldsymbol B}}
%\newcommand{\grad}{\ensuremath{\boldsymbol \nabla}}
%\newcommand{\gradp}{\ensuremath{\boldsymbol \nabla}_p}
%\newcommand{\ppos}{\ensuremath{\boldsymbol x}}

@@

%\newcommand{\betavec}{\ensuremath{\boldsymbol \beta}}
%\newcommand{\betaLAB}{\ensuremath{\boldsymbol \beta_{\rm L}}}
%\newcommand{\betaCM}{\ensuremath{\boldsymbol \beta_{\rm COM}}}
\newcommand{\betaCMpara}{\ensuremath{\beta_{\rm COM,\parallel}}}
\newcommand{\betaCMperp}{\ensuremath{\beta_{\rm COM,\perp}}}

@@

%\newcommand{\betabar}{\ensuremath{\overline{\boldsymbol \beta}}}
\newcommand{\betabar}{\ensuremath{\overline{\bmath{\beta}}}}
\newcommand{\betabarLAB}{\ensuremath{\betabar_{\rm L}}}
\newcommand{\betabarCM}{\ensuremath{\betabar_{\rm COM}}}
\newcommand{\betaBOOST}{\ensuremath{\beta_b}}

@@

%\newcommand{\kvec}{\ensuremath{\boldsymbol k}}
%\newcommand{\rvec}{\ensuremath{\boldsymbol r}}
%\newcommand{\rvec}{\ensuremath{\boldsymbol j}}

@@

%\newcommand{\xhat}{\ensuremath{\hat{\boldsymbol x}}}
%\newcommand{\yhat}{\ensuremath{\hat{\boldsymbol y}}}
%\newcommand{\zhat}{\ensuremath{\hat{\boldsymbol z}}}

@@

\newcommand{\betabeam}{\ensuremath{\bmath{\beta}_{\rm b}}}
\newcommand{\betabeamz}{\ensuremath{\beta_{\rm b,z}}}
\newcommand{\betabeamx}{\ensuremath{\beta_{\rm b,x}}}
\newcommand{\gammabeam}{\ensuremath{\gamma_{\rm b}}}
\newcommand{\vbeam}{\ensuremath{v_{\rm b}}}
\newcommand{\gammaphase}{\ensuremath{\gamma_{\rm ph}}}
\newcommand{\betaphase}{\ensuremath{\beta_{\rm ph}}}

@@

%\author{Some people}
\author{
Philip Chang\altaffilmark{1},
Avery E.~Broderick\altaffilmark{2,3},
Christoph Pfrommer\altaffilmark{4},
Ewald Puchwein\altaffilmark{5},
Astrid Lamberts\altaffilmark{6},
Mohamad Shalaby\altaffilmark{2,3,7},
Geoffrey Vasil\altaffilmark{8}
}
\altaffiltext{1}{Department of Physics, University of Wisconsin-Milwaukee, 3135 North Maryland Avenue, Milwaukee, WI 53211, USA; chang65@uwm.edu}
\altaffiltext{2}{Perimeter Institute for Theoretical Physics, 31 Caroline Street North, Waterloo, ON, N2L 2Y5, Canada}
\altaffiltext{3}{Department of Physics and Astronomy, University of Waterloo, 200 University Avenue West, Waterloo, ON, N2L 3G1, Canada}
\altaffiltext{4}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}
\altaffiltext{5}{Institute of Astronomy and Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
\altaffiltext{6}{TAPIR, Mailcode 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\altaffiltext{7}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
\altaffiltext{8}{School of Mathematics \& Statistics, University of Sydney, NSW 2006, Australia}

@@

%\maketitle

\begin{abstract}
The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that they drive.  Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes.  We reproduce the well-known reactive and kinetic growth rates for both the beam plasma and oblique mode. We demonstrate for the oblique instability that there is broad spectrum of unstable modes that grow at the maximum rate for a wide range of beam temperatures and wavevector orientations relative to the beam.  We also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower Lorentz factor than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.  Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the both the kinetic oblique mode and the reactive beam-plasma mode.  These results are important in understanding how powerful plasma instabilities may sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space.
%\ab{[The only issue with the above is that the primary issues people have raised are with nonlinear effects, e.g., saturation or nonlinear coupling to existing background structures.  Should we be quite as strident then in the concluding sentences?]}
\end{abstract}

@@

%CP: this appears twice - I cut the following:
%This excess heating of the IGM may resolve a variety
%of cosmological puzzles, including naturally explaining
%  anomalies in the statistics of the high-redshift Ly$\alpha$ forest
%\citep{paperIV} and potentially explaining a number of the X-ray
%  properties of groups and clusters and anomalies in galaxy formation on
%  the scale of dwarfs \citep{paperIII,Lu+2013}.  We
%  have recently shown that \textit{if} the IC halos are ignored, it is possible
%  to quantitatively reproduce the redshift and flux distributions
%  of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
%  spectrum above 3~GeV simultaneously with a unified model of AGN evolution
%  \citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
%  circumstantial evidence for the presence of virulent plasma beam
%  instabilities.

The \Fermi satellite and ground-based imaging atmospheric Cherenkov telescopes such as
% Phil: I added "imaging atmospheric" to distinguish this class of
% telescopes from imaging water CTs, e.g., MILAGRO; I am not aware of
% any space based IACT apart from a planned telescope on the ISS, but
% who knows whether this will ever fly
H.E.S.S., MAGIC, and VERITAS\footnote{High Energy
  Stereoscopic System, Major Atmospheric Gamma Imaging Cerenkov Telescope, Very
  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the high
energy Universe is teeming with energetic very high-energy gamma-ray 
(VHEGR, $E > 100\,\GeV$) sources, the extragalactic component of which mainly consists of TeV
blazars with a minority population of other sources  such as radio
and starburst galaxies. These extragalactic VHEGR emitters produce TeV photons that are greatly attenuated via annihilation upon soft photons in the extragalactic background light
(EBL) and produce pairs \citep[see, e.g.,][]{Goul-Schr:67,Sala-Stec:98,Nero-Semi:09}.

@@

These two important conclusions depend on IC cooling dominating the evolution of the ultra-relativistic pairs.  However, it was recently found that plasma instabilities driven by the
ultrarelativistic pair beams  likely are the dominant cooling mechanisms
(\citealt{paperI}, hereafter BCP12, \citealt{Schlickeiser+12,Schlickeiser+13}), depositing this energy
as heat in the intergalactic medium \citep{paperII,paperIII}.  Therefore, the lack of an observed
 IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
(BCP12;\citealt{Schlickeiser+12,Schlickeiser+13}).

% Phil: I cut this part because it is written as a fact (while it is not) and the same long author list is repeated in the next paragraph (the \qc command did not work with the \citep command):
% and intergalactic magnetic fields (IGMF)
% \citep[IGMF; ][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.

@@

@@

% \ac{[There are some notational issues throughout the text, which I'll list here.  Usually it is simply a matter of choosing, and then sticking to, a convention.
% \begin{itemize}
% \item Some equations assume $c=1$ while others have the $c$'s explicitly present.  Unless there is a compelling reason (e.g., clarity of derivation vs. utility of an answer in physical units) we should choose one and mention the convention here (at the end of the intro) if $c=1$.
% \item $\beta$ and $v$ appear to be used interchangeably, again related to the $c$ or not question.  Again, we should choose one.
% \item Throughout section 2 and in the Appendix the frame is defined in terms of $\parallel$ and $\perp$ directions (i.e., $\beta_\parallel$ and $\bmath{\beta}_\perp$, note the latter is a {\em vector}!) while in sections 3 and in the figures we refer to $\beta_z$ and $\beta_x$ for these.
% \item ApJ style, which I'm assuming is what we are following, has vectors given by bold, roman fonts and scalars by the italic fonts.  Thus, it should be $\bmath{k}\cdot\bmath{\delta E}$ or $\bmath{k}\cdot\delta\bmath{E}$, etc.  I've tried to fix up various definitions in the preamble to do this, but that should be checked.
% \end{itemize}
% In any equations I've added/modified I will include the $c$'s and use $\beta$ for concreteness, though I anticipate there will be few such examples.]}

@@

As stated in the Introduction, VHEGR photons pair produce on encountering EBL
photons as they propagate throughout the universe \citep{Gould+66}, and this
attenuation of VHEGR flux has been used as a probe of the EBL
\citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}.  The basic requirement
of this process is that the energies of the VHEGR ($E_{\rm ph}$) and the EBL
photon ($E_{\rm ebl}$) exceed the rest mass energy of the $\epm$ pair in the
center of momentum (COM \ab{[Should we use CM instead for compactness? Otherwise ``of'' is getting a big promotion!]}) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
frame. As a result, an $\epm$ pair can be produced with Lorentz factor$\gamma = \left(1-v^2/c^2\right)^{-1/2} \approx E/2m_e c^2$, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
pair beam that emerges from this process.

%As the mean free path of these VHEGR photons to pair production is short compared to the Hubble length, observations of extragalactic VHEGR sources have been used as probes of the EBL \citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}

@@

The relativistic Maxwellian distribution can be extended to a drifting (or boosted) distribution via an appropriate Lorentz transformation.  The relationship between the energies of the lab (boosted) frame and the COM frame is
\begin{equation}
  E_{\rm COM} = \gammabeam\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right),
\end{equation}
where $\gammabeam = \gamma(v_{\rm b}) = \left(1 - v_{\rm b}^2/c^2\right)$ is the Lorentz factor of the beam and $v_{\rm b}$ is the bulk velocity of the pair beam.
Inserting this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
\begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
%f = \frac{1}{4\pi\gammabeam m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
f = \frac{n_b m_e c^2}{4\pi\gammabeam k_B T_b K_2(m_ec^2/k_B T_b )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - \vel_b\cdot \bmath{p})} {k_B T_b}\right),
\end{eqnarray}
where $K_2$ is the 2nd order modified Bessel function and $T_b$ is the comoving temperature of the beam.

%\ab{[Do we want to describe the characteristics of this distribution at any further length?  I recall you had a detailed discussion of the characteristics of a boosted thermal distribution which was nice because it clarified some obvious misconceptions that seemed to continuously arise.  Also, do we want to discuss the potential evolution of the distribution?]}

The Maxwell-J{\"u}ttner distribution leads to an anisotropic velocity spread parallel and perpendicular to the beam's direction.  In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale.  The relevant results are: 
\begin{equation}\label{eq:perp dispersion}
\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 }%\sout{,}
\quad\text{and}\quad
\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
\end{equation}
%and 
%\begin{equation}\label{eq:parallel dispersion}
%\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
%\end{equation}
where $T$ is measured in the COM frame of the beam.  
These simple scalings of the perpendicular velocity dispersion and parallel velocity dispersions can be understood as a results of time dilation between two frames that are boosted relative to each other, giving one factor of $\gamma^{-1}$. The coordinates perpendicular to the boost axis remain invariant while the axis along the boost suffers from length contraction and gives an extra scaling of $\gamma^{-1}$ for the parallel case.  In any case, an ultrarelativistic beam has a small velocity spread in both the parallel and perpendicular directions by factors of $\gamma^{-2}$ and $\gamma^{-1}$, respectively.  These velocity dispersions will be important in delineating the regime of instability in \S~\ref{sec:transition}.

%The transformation properties of the drifting Maxwell-J{\"u}ttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-J{\"u}ttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).

%In Appendix~\ref{sec:lorentz}, we explicitly derive the scaling of the
%  parallel and perpendicular velocity dispersions with the Lorentz factor upon
%  boosting the distribution function to the lab frame.
%In the limit in which the perpendicular velocity dispersion in the COM frame is non-relativistic, i.e.,
%$\left|\betaCM\right| \ll 1$, we find to lowest order in $\betaCM$
%\begin{equation}
% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2}.
% \label{eq:betaperp}
%\end{equation}
%This simple scaling of the perpendicular velocity dispersion is an immediate 
%consequence of how transverse velocities transform under Lorentz boosts, and may be understood
%in terms of time dilation.
%
%This result is also anticipated by the standard transformation of temperature under boosts, $T\rightarrow T/\gamma$ (see Equation~\ref{eq:drifting maxwellian}).
%This can be understood by considering $mv^2 \sim kT$ -- we pick up two factors of $1/\gamma$
%from the perpendicular velocity dispersion that is countered by one factor of
%$\gamma$ from the mass scaling. 
%
%For the parallel component of the velocity
%dispersion in the non-relativistic limit, $\left|\betaCM\right| \ll 1$, we find
%to lowest order in $\betaCM$:
%\begin{equation}
% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gammabeam^{4}},
% \label{eq:betapara}
%\end{equation}
%Again this scaling follows immediately from the transformation of velocities under Lorentz boosts, and may be simply understood as a consequence of both time dilation and, for the parallel case, length contraction.
%This scaling of the parallel component of the velocity
%dispersion has important consequences below.
%

% To see how this distribution function appears in velocity space, we show four examples of the  Maxwell-J{\"u}ttner distribution in the center-of-mass frame for the nonrelativistic temperatures ($T=10^{-2}$ and $ 0.1$) in Figure \ref{fig:distribution_non_rel} and for relativistic temperatures ( $T=1$ and $10$) in Figure \ref{fig:distribution_rel}, where $T$ is listed in the natural relativistic units: $m_e c^2/k_B$.
% We plot these in terms of $\beta_z$, which is parallel to the beam and $\beta_x$, which is perpendicular to the beam direction though, in the center-of-mass frame, an isotropic distribution makes no distinction between these directions.  We note that not all combinations of $\beta_x$ and $\beta_z$ are allowed as $\beta_x^2 + \beta_z^2 < 1$.  \ac{[Could you blank out the portions for which $\beta^2>=1$ in the figure to both drive this point home as well has emphasize the scaling difference?]}
% 
% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gammabeam = 10^6$.  Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gammabeam^{-1}$ and the parallel component by $\gammabeam^{-2}$.
% When $T\ll1$, as anticipated above, the parallel and transverse velocity dispersions are well approximated by the scaling in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}).  As seen explicitly in Figure \ref{fig:distribution_non_rel}, the distribution functions remain roughly isotropic in the rescaled and shifted velocity units.
% 
% For relativistic temperatures, the conditions under which the expressions in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}) were derived no longer hold.  Nevertheless, despite the importance of higher-order and anisotropic terms in the distribution function, the magnitudes of the velocity dispersions remain well approximated by these scalings.  This may be seen explicitly in the right panels of Figure \ref{fig:distribution_rel}.
% While these scaling will necessarily break down as $T \rightarrow \gammabeam$, for the problem of interest this is never the case.  Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gammabeam^{-1}$ in the perpendicular direction and like $\gammabeam^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gammabeam \gg T$).  
% 
% \begin{figure*}
%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_0_01.pdf}
%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_0_01.pdf}\\
%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_0_1.pdf}
%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_0_1.pdf}\\
%  \caption{Distribution function in velocity space in the center-of-mass frame (left) and the boosted (lab) frame (right) for nonrelativistic temperatures: $T=0.01$ (top) and $0.1$ (bottom).  Note that while we plot $\beta_z$ and $\beta_x$ from -1 to 1, not all values of $\beta_z$ and $\beta_x$ are allowed. \ac{[Can this be shown explicitly, e.g., by making those values NaN, or some other method that results in no plot in that instance?  Also, how are these normalized?]}  On the right plot, we show the distribution in a boosted frame by a Lorentz factor of $10^6$.  As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_b$ on the x-axis.  We have also multiplied the $\beta$ factors by a factor of $\gammabeam^2$ for the z (parallel) component and $\gammabeam$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}. These plots show that a non-relativistic velocity dispersion transforms as $\gammabeam^{-2}$ and $\gammabeam^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}
% \end{figure*}
% \begin{figure*}
% \includegraphics[width=0.5\textwidth]{distribution_cm_E_1.pdf}
%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_1.pdf}\\
%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_10.pdf}

>>>>

%\documentclass[usenatbib,iop,apj]{emulateapj}
%\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2015}
\documentclass[usenatbib,iop,apj,numberedappendix]{aastex}

@@

\newcommand{\mathsc}[1]{\text{\textsc{#1}}}

@@

@@

\newcommand{\betaCMpara}{\ensuremath{\beta_{\rm COM,\parallel}}}
\newcommand{\betaCMperp}{\ensuremath{\beta_{\rm COM,\perp}}}

@@

\newcommand{\betabar}{\ensuremath{\overline{\bmath{\beta}}}}
\newcommand{\betabarLAB}{\ensuremath{\betabar_{\rm L}}}
\newcommand{\betabarCM}{\ensuremath{\betabar_{\rm COM}}}
\newcommand{\betaBOOST}{\ensuremath{\beta_b}}

@@

@@

@@

\newcommand{\betabeam}{\ensuremath{\bmath{\beta}_{b}}}
\newcommand{\betabeamz}{\ensuremath{\beta_{b,z}}}
\newcommand{\betabeamx}{\ensuremath{\beta_{b,x}}}
\newcommand{\gammabeam}{\ensuremath{\gamma_{b}}}
\newcommand{\vbeam}{\ensuremath{v_{b}}}
\newcommand{\gammaphase}{\ensuremath{\gamma_{\rm ph}}}
\newcommand{\betaphase}{\ensuremath{\beta_{\rm ph}}}

@@

\author{
Philip Chang\altaffilmark{1},
Avery E.~Broderick\altaffilmark{2,3},
Christoph Pfrommer\altaffilmark{4},
Ewald Puchwein\altaffilmark{5},
Astrid Lamberts\altaffilmark{6},
Mohamad Shalaby\altaffilmark{2,3,7},
Geoffrey Vasil\altaffilmark{8}
}
\altaffiltext{1}{Department of Physics, University of Wisconsin-Milwaukee, 3135 North Maryland Avenue, Milwaukee, WI 53211, USA; chang65@uwm.edu}
\altaffiltext{2}{Perimeter Institute for Theoretical Physics, 31 Caroline Street North, Waterloo, ON, N2L 2Y5, Canada}
\altaffiltext{3}{Department of Physics and Astronomy, University of Waterloo, 200 University Avenue West, Waterloo, ON, N2L 3G1, Canada}
\altaffiltext{4}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}
\altaffiltext{5}{Institute of Astronomy and Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
\altaffiltext{6}{TAPIR, Mailcode 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
\altaffiltext{7}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
\altaffiltext{8}{School of Mathematics \& Statistics, University of Sydney, NSW 2006, Australia}

@@

\begin{abstract}
The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that they drive.  Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes.  We reproduce the well-known reactive and kinetic growth rates for both the beam plasma and oblique mode. We demonstrate for the oblique instability that there is broad spectrum of unstable modes that grow at the maximum rate for a wide range of beam temperatures and wavevector orientations relative to the beam.  We also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower Lorentz factor than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.  Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the both the kinetic oblique mode and the reactive beam-plasma mode.  These results are important in understanding how powerful plasma instabilities may sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space.
\end{abstract}

@@

The \Fermi satellite and ground-based imaging atmospheric Cherenkov telescopes such as
H.E.S.S., MAGIC, and VERITAS\footnote{High Energy
  Stereoscopic System, Major Atmospheric Gamma Imaging Cerenkov Telescope, Very
  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the high
energy Universe is teeming with energetic very high-energy gamma-ray 
(VHEGR, $E > 100\,\GeV$) sources, the extragalactic component of which mainly consists of TeV
blazars with a minority population of other sources  such as radio
and starburst galaxies. These extragalactic VHEGR emitters produce TeV photons that are greatly attenuated via annihilation upon soft photons in the extragalactic background light
(EBL) and produce pairs \citep[see, e.g.,][]{Goul-Schr:67,Sala-Stec:98,Nero-Semi:09}.

@@

These two important conclusions depend on IC cooling dominating the evolution of the ultra-relativistic pairs.  However, it was recently found that plasma instabilities driven by the
ultrarelativistic pair beams  likely are the dominant cooling mechanisms
(\citealt{paperI}, hereafter BCP12, \citealt{Schlickeiser+12,Schlickeiser+13}), depositing this energy
as heat in the intergalactic medium \citep{paperII,paperIII}.  Therefore, the lack of an observed
 IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
(BCP12;\citealt{Schlickeiser+12,Schlickeiser+13}). {\bf We note that this effectiveness of this plasma instabilities is complicated by nonlinear effects, which we briefly discuss below.}

@@

{\bf Additional nonlinear effects may also be important. For instance, for sufficiently powerful blazers, the modulation instability may operate \citep{Schlickeiser+12,Chang+14,2015MNRAS.448.3405M}, allowing for a rapid transfer of electrostatic wave energy into thermal energy.  For less powerful blazars, the combination of NLD and quasilinear damping, i.e., beam plateauing, will also reduce the rate of damping compared to the linear rate, and alters the resulting IC spectra \citep{2015MNRAS.448.3405M}. Further study of these effects will help clarify these points.}

@@

@@

As stated in the Introduction, VHEGR photons pair produce on encountering EBL
photons as they propagate throughout the universe \citep{Gould+66}, and this
attenuation of VHEGR flux has been used as a probe of the EBL
\citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}.  The basic requirement
of this process is that the energies of the VHEGR ($E_{\rm ph}$) and the EBL
photon ($E_{\rm ebl}$) exceed the rest mass energy of the $\epm$ pair in the
center of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
frame. As a result, an $\epm$ pair can be produced with Lorentz factor$\gamma = \left(1-v^2/c^2\right)^{-1/2} \approx E/2m_e c^2$, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
pair beam that emerges from this process.

@@

The relativistic Maxwellian distribution can be extended to a drifting (or boosted) distribution via an appropriate Lorentz transformation.  The relationship between the energies of the lab (boosted) frame and the COM frame is
\begin{equation}
  E_{\rm COM} = \gammabeam\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right),
\end{equation}
where $\gammabeam = \gamma(v_b) = \left(1 - v_b^2/c^2\right)$ is the Lorentz factor of the beam and $v_b$ is the bulk velocity of the pair beam.
Inserting this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
\begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
f = \frac{n_b m_e c^2}{4\pi\gammabeam k_B T_b K_2(m_ec^2/k_B T_b )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - \vel_b\cdot \bmath{p})} {k_B T_b}\right),
\end{eqnarray}
where $K_2$ is the 2nd order modified Bessel function and $T_b$ is the comoving temperature of the beam.

The Maxwell-J{\"u}ttner distribution leads to an anisotropic velocity spread parallel and perpendicular to the beam's direction.  In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale.  The relevant results are: 
\begin{equation}\label{eq:perp dispersion}
\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 }%\sout{,}
\quad\text{and}\quad
\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
\end{equation}
where $T$ is measured in the COM frame of the beam.  
These simple scalings of the perpendicular velocity dispersion and parallel velocity dispersions can be understood as a results of time dilation between two frames that are boosted relative to each other, giving one factor of $\gamma^{-1}$. The coordinates perpendicular to the boost axis remain invariant while the axis along the boost suffers from length contraction and gives an extra scaling of $\gamma^{-1}$ for the parallel case.  In any case, an ultrarelativistic beam has a small velocity spread in both the parallel and perpendicular directions by factors of $\gamma^{-2}$ and $\gamma^{-1}$, respectively.  These velocity dispersions will be important in delineating the regime of instability in \S~\ref{sec:transition}.

{\bf While we have modeled the pair distribution function as a Maxwell-J{\"u}ttner distribution, the distribution function that are produced by VHEGR photo annihilation is different \citep[see for instance][]{2012ApJ...758..101S}.  In particular, the the parallel and perpendicular momentum spread will likely be different due to the distribution of VHEGR photons and their respective mean free paths.  However, a Maxwell-J{\"u}ttner distribution is still useful as it has been calculated without approximation previously by \citet{Bret-Grem-Beni:10}, allowing a point of comparison for our calculation using the electrostatic approximation (as mentioned below).  In addition, it possess a continuous (small) distribution of parallel and perpendicular momenta that allow us to elucidate the physics.  It also is sufficiently simple to allows us to calculate the kinetic instability exactly in the electrostatic approximation.  Finally, we note that the analytic methodology use to calculate the Maxwell-J{\"u}ttner distribution may be useful for the full calculation using the physical distribution function.}
